---
title: "Data Wrangling in R! It's Super Effective!"
output: learnr::tutorial
runtime: shiny_prerendered
subtitle: "An Introduction to Data Wrangling with dplyr"
author: Madison Corinha & Catherine Ciulla
---

```{r setup, include=FALSE}
library(learnr)
library(readr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
pokemon <- read_csv("pokemon.csv")
```

## Introduction

Hello!

In your Data Science journey so far, you've charted new territory in Data Exploration, you've seen the sights in Data Visualization, so what next? 

Put on your cowboy boots and wide-brimmed hats, because we're about to go Data Wranglin'.


In this tutorial we will cover some of the basics of Data Wrangling using the `dplyr` function in the `tidyverse` package.

### What is Data Wrangling?

Data Wrangling is the process of gathering, selecting, cleaning, structuring, transforming, and enriching "raw" data to turn it into a format that is more appropriate and convenient to use.

### Data

In this tutorial, we will use the `pokemon.csv` dataset uploaded by Rounak Banik, which you can find [here](https://www.kaggle.com/rounakbanik/pokemon).

The Kaggle page where this dataset was posted described the data as follows:

> This dataset contains information on all 802 Pokemon from all Seven Generations of Pokemon. The information contained in this dataset include Base Stats, Performance against Other Types, Height, Weight, Classification, Egg Steps, Experience Points, Abilities, etc. The information was scraped from http://serebii.net/


## 








## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

