---
title: "Data Wrangling in R! It's Super Effective!"
output: learnr::tutorial
runtime: shiny_prerendered
subtitle: "An Introduction to Data Wrangling with dplyr"
author: Madison Corinha & Catherine Ciulla
---

```{r setup, include=FALSE}
library(learnr)
library(readr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
pokemon <- read_csv("pokemon.csv")
```


## Welcome!

In your Data Science journey so far, you've charted new territory in Data Exploration, you've seen the sights in Data Visualization, so what next? 

Put on your cowboy boots and wide-brimmed hats, because we're about to go Data Wranglin'.


In this tutorial we will cover some of the basics of Data Wrangling using the `dplyr` function in the `tidyverse` package.

### What is Data Wrangling?

Data Wrangling is the process of gathering, selecting, cleaning, structuring, transforming, and enriching "raw" data to turn it into a format that is more appropriate and convenient to use.

### Data

In this tutorial, we will use the `pokemon.csv` dataset uploaded by Rounak Banik, which you can find [here](https://www.kaggle.com/rounakbanik/pokemon).

The Kaggle page where this dataset was posted described the data as follows:

> This dataset contains information on all 802 Pokemon from all Seven Generations of Pokemon. The information contained in this dataset include Base Stats, Performance against Other Types, Height, Weight, Classification, Egg Steps, Experience Points, Abilities, etc. The information was scraped from http://serebii.net/


## Introduction

### Verb-Named Functions

The functions in the `dplyr` package are named with the verbs that describe the action being performed by the function. 

For example:
  - The `arrange()` function *arranges* (sorts) the rows of the dataset 
  - The `select()` function *selects* a subset of variables (columns) from the dataset

For this reason, the `dplyr` package provides a **grammar** of data manipulation.

### Pipelines with %>% 

The verb-named functions in the `dplyr` package aren't meant to only work alone, they can also be used in conjunction with each other in a string of functions called a pipeline. 

In a pipeline, multiple arguments are linked together with the **pipe operator**, **%>%**. In a pipeline, the arguments at the top (on the left) of the pipe operator are performed first, and then the argument on the bottom (on the right). Multiple arguments can be piped together to create a long string of actions.

To start, lets use a pretty universal example: Let's make a hypothetical pipeline (without code) that would show the process of drinking some water. The pipeline would look something like this:

  - "get cup" %>% "fill cup with water" %>% "bring cup to mouth" %>% "drink the water"
  
The pipeline starts off with grabbing a cup, and when that action is completed, the next action begins - in this case, its filling the cup with water. When the cup is filled with water, the next action happens, and after that, the next action happens. Each step in the pipeline must be completed before starting the next step - you can't fill a cup with water if you haven't grabbed the cup yet. At the end of the pipeline, you have completed the action of drinking water. 

In the following sections, you might come across practice questions like the one shown below. Use this question to practice your knowledge of pipelines.
```{r practice_pipeline_question}

question("Fill in the [...] in the pipeline if the goal is to catch a pokemon: 'walk into tall grass' %>%  'encounter wild pokemon' %>% [...] %>% 'capture pokemon' %>%  'pick up Pokeball'",
         answer("Introduce yourself and shake hands like professional adults"),
         answer("Complain about the Boston Red Sox trading Mookie Betts"),
         answer("Throw a Pokeball", correct = TRUE),
         answer("Contemplate the ethics of taking a creature out of its natural environment and forcing it to fight")
         )

```

Let's now use an example that's more closely related to what we are going to be using pipelines for in this tutorial. When using a pipeline for data wrangling purposes, it almost always follows the same basic steps. The first step in the pipe is to specify the name of the dataset that you are performing the actions on, then the following steps are the actions you want to perform on that dataset.

  - dataset_name %>% action(...) %>% action(...) %>% action(...)

Just like you have to get the cup before you can do anything to it, you have to specify the dataset name before you can do anything to it.



## select()

The `select()` function returns to you a subset of columns. In a dataset, the columns represent the *variables*.

This function is helpful for isolating and analyzing specific variables in a dataset. 

### Practice

The `select()` function only returns the variables you tell it to return from a dataset.

The original dimensions of the `pokemon` dataset are very large! There are 801 cases (rows) and 41 variables (columns)! We don't need the information from all of those variables for the purposes of this tutorial, so let's be sure to use `select()` so we only see the variables we want.

Let's use the `select()` function to only return the english name of the pokemon (`name`), the type of the pokemon (`type1`), and the generation the pokemon first appeared (`generation`):
```{r Example1, exercise = TRUE, exercise.eval = TRUE}
pokemon %>% 
  select(name, type1, generation)
```

Now you try: 

Use the `select()` function to return the English name of the pokemon (`name`), the type of the pokemon (`type1`), the health of the pokemon (`hp`), and the weight in kilograms (`weight_kg`). Use the "Hint" button if you need help.
```{r Exercise2, exercise = TRUE, warning=FALSE}

```

```{r Exercise2-hint}
pokemon %>% 
  select(name, type1, hp, weight_kg)
```

Now that you've got the hang of `select()`, let's move on to `filter()`.

## filter()

The `filter()` function returns to you a subset of rows. In a dataset, the rows represent the *cases, observations, or individuals*.

This function allows you to specify criteria about values in a dataset by `filter()`ing through them to return only the rows that fit what you're looking for. 

The `filter()` and `select()` functions can be used together in a single pipeline in order to isolate specific variables *and* values. You can also `filter()` multiple criteria with the `&` operator.

### Practice

I want to find only the pokemon with really high amounts of base health points (`hp`). Let's `filter()` the dataset to only include pokemon with greater than 100 base HP, and only return the variables for English name (`name`), Health Points (`hp`), and type (`type1`).

```{r exercise3, exercise = TRUE, exercise.eval = TRUE}
pokemon %>% 
  select(name, hp, type1) %>% 
  filter(hp > 100)
```

It is good to note that filtering for `hp > 100` does NOT organize the results into ascending/descending order!

What if we wanted to `filter()` by a categorical variable like pokemon type (`type1`) instead of a numerical variable like Health points (`hp`)? We can do that! Even better, we can do both at the same time! (Note: Since the `=` operator is used for *assignment*, we use the `==` operator to *test for equality*)

Let's use the same code as before but add another layer to the `filter()` argument: return *only* water type pokemon *that also have* base Health Points greater than 100.
```{r exercise4, exercise = TRUE, exercise.eval = TRUE}
pokemon %>% 
  select(name, hp, type1) %>% 
  filter(hp > 100 & type1 == "water")
```

Important to note, again, that the list is NOT in an organized order, it is just in the order they happened to be listed in the original dataset.

This time, re-use the code from above and change it to `filter()` for any pokemon type that you like. 
If you want it, you can find information about pokemon types [here](https://bulbapedia.bulbagarden.net/wiki/Type).

```{r exercise5, exercise = TRUE, exercise.lines = 5}

```

Hint: If no pokemon show up in the list for your specific type, try changing the number for `hp > ...` to something lower.

## mutate()

The `mutate()` function allows you to create and insert new columns (variables) into a dataset. 

You can use algebraic math equations to define the new variables (multiplying, dividing, etc.)

### Practice

Pokemon have different values for general attack power and for special attack power. Let's use the `mutate()` function to combine these two values into a new variable called `base_total_offense`.

```{r exercise6, exercise = TRUE, exercise.eval = TRUE}
pokemon %>% 
  mutate(base_total_offense = attack + sp_attack) %>% 
  select(name, attack, sp_attack, base_total_offense)
```

**Note**: notice how we used `select()` *after* we used the `mutate()` function. Why do you think that was necessary? (Hint: you can't fill a cup with water if you haven't grabbed a cup yet!) 

## arrange()

The `arrange()` function organizes the values of a dataframe based on a specific variable. 

`arrange()` sorts in ascending order automatically, but you can use `desc()` to sort in descending order for any of the variables specified in `arrange()`. 

### Practice

Remember when we pointed out that the `filter()` function didn't change the order of the data? We can fix that by using the `arrange()` function!

Let's use the same code from the first exercise in the `filter()` section, and use `arrange()` to put it in order of greatest to least HP.
```{r exercise7, exercise = TRUE, exercise.eval = TRUE}
pokemon %>% 
  select(name, hp, type1) %>% 
  filter(hp > 100) %>% 
  arrange(desc(hp))
```

Looks like the pokemon Blissey has a crazy high base HP! 

```{r arrange_mc_question}
question("If I wanted to `arrange()` the data to sort by *least to greatest* attack power, which of the options below would I use?",
         answer("arrange(attack)", correct = TRUE),
         answer("arrange(desc(attack))"),
         answer("arrange(least_to_greatest)")
         )
```


## summarize() with group_by()

The `summarize()` and `group_by()` functions let us make calculations and comparisons *within* the data.

The `summarize()` function collapses the values into rows based on defined summary statistics. We can use functions such as `median()`, `mean()`, `sd()`, etc. to specify the summary statistic we want to use. 

The `summarize()` function is *almost always* used in conjunction with the `group_by()` function. 

We can use the `group_by()` function to generate aggregate summary results. It is important to keep in mind that the `group_by()` function does nothing by itself, and thus should never be used at the end of a pipeline!

### Practice

In this example, we are grouping by pokemon type, so all data about pokemon of the same type will be considered together as a group. With this, we are using `summarize()` to calculate the mean weight (in kg) of all pokemon in each of the different pokemon types. We are then using the `arrange()` function to sort the results in descending order.

```{r exercise8, exercise = TRUE, exercise.eval = TRUE}
pokemon %>%
  group_by(type1) %>%
  summarize(mean_weight = mean(weight_kg, na.rm = TRUE)) %>% # "na.rm = TRUE" removes all missing values so they don't mess up the calculations! 
  arrange(desc(mean_weight))

```

Now you try:

Change this code so that we are calculating average height (`height_m`) grouped by type (`type1`). Use the "hint" button to check your work.

```{r exercise9, exercise = TRUE}
pokemon %>%
  group_by(type1) %>%
  summarize(mean_height = mean(    , na.rm = TRUE)) %>%
  arrange(desc(   ))
```

```{r exercise9-hint}
pokemon %>%
  group_by(type1) %>%
  summarize(mean_height = mean(height_m, na.rm = TRUE)) %>%
  arrange(desc(mean_height))
```



## slice()

`slice()` allows you to pick rows by their original position in the table (ex. `slice(1:10)` returns rows 1 through 10)



## count()

`count()` is the equivalent of `group_by()` + `tally()`.

If data is already grouped, `count()` adds an additional group that is removed afterwards.

### Practice

Is there a trend in the number of new pokemon introduced in each generation?

```{r exercise10, exercise = TRUE, exercise.eval = TRUE}
pokemon %>%
  count(generation, sort = TRUE, name = "n_pokemon_per_gen") %>% 
  arrange(generation)

```







********



## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

